---
output: html_document
---
# Simulation Exercise for the Exponential Distribution
## Author: Edward Lau

### Overview: I am going to report on a simulation exercise and inferential data analysis. The first part will examine how the exponential distribution behaves under simulation. The second part will analyze the ToothGrowth data in the R dataset package.

## Part 1: Exponential distribution simulation

The following code will do 1000 simulations of random exponential distribution with lambda = 0.2
```{r}
sim <- 1000
means = NULL
vars = NULL
for (i in 1: sim){
        means <- c(means, mean(rexp(40,0.2)))
        vars <- c(vars, var(rexp(40,0.2)))
        }
```

The histogram of the simulation result is below:
```{r}
hist(means)
```

###Question 1: The mean of the simulated values:
```{r}
mean(means)
```
which compares quite well to the theoretical value of 1/0.2 = 5.

###Question 2: The standard deviation of the distribution:
```{r}
hist(vars^0.5)
mean(vars^0.5)
```
Hence the variance of the simulation is about (1/0.2)^2 ~ 25; and the standard deviation ~5, which compares quite well to the theoretical value of 1/0.2 = 5.

###Question 3: Normal distribution?
```{r}
quantile(means)
quantile(rnorm(1000,mean=mean(means), sd=sd(means)))
```
It seems like the quantiles of the two distributions are quite comparable...

## Part 2: Inference
This will load the ToothGrowth dataset, which examines the effect of Vitamin C on tooth growth in guinea pigs. len = length of teeth for 10 guinea pigs; supp = OC or VC delivery methods (orange juice or ascorbic acid); dose = dose level

###Question 1: Load the ToothGrowth data and perform basic exploratory data analyses
```{r}
library(datasets)
library(dplyr)
library(ggplot2)
glimpse(ToothGrowth)
qplot(len, dose, data = ToothGrowth, facets = .~supp)
```
Conclusion: it looks like there are increases in tooth length by doses of OJ and VC. 

###Question 2: Summarizing the data by dose and supp.
```{r}
group_by(ToothGrowth, supp, dose) %>% summarize(mean = mean(len))
```

###Question 3: Compare dose and supplement delivery method using a t test
```{r}
OJdata_0.5 <-filter(ToothGrowth, supp=="OJ", dose == 0.5)
VCdata_0.5 <-filter(ToothGrowth, supp=="VC", dose == 0.5)
t.test(OJdata_0.5$len,VCdata_0.5$len)

OJdata_1.0 <-filter(ToothGrowth, supp=="OJ", dose == 1.0)
VCdata_1.0 <-filter(ToothGrowth, supp=="VC", dose == 1.0)
t.test(OJdata_1.0$len,VCdata_1.0$len)

OJdata_2.0 <-filter(ToothGrowth, supp=="OJ", dose == 2.0)
VCdata_2.0 <-filter(ToothGrowth, supp=="VC", dose == 2.0)
t.test(OJdata_2.0$len,VCdata_2.0$len)
```
Conclusion: it appears that there is a significant difference in tooth length when given low doses (0.5 and 1.0) of vitamin C through orange juice and through ascorbic acids, but this significance is no longer there at high doses (2.0)

