---
output: word_document
---
---
output: html_document
--- Inference on ToothGrowth
## Author: Edward Lau

### Overview: I am going to perform statistical inference on the ToothGrowth dataset in the R datasets package, to ask the question of whether the dose and delivery methods of vitamin C may have a significant effect on the growth of odontoblasts (teeth) in a population of 10 guinea pigs.

### Part 2: Inference
This will load the ToothGrowth dataset, which examines the effect of Vitamin C on tooth growth in guinea pigs. len = length of teeth for 10 guinea pigs; supp = OC or VC delivery methods (orange juice or ascorbic acid); dose = dose level

###Question 1: Load the ToothGrowth data and perform basic exploratory data analyses
```{r}
library(datasets)
library(dplyr)
library(ggplot2)
glimpse(ToothGrowth)
qplot(len, dose, data = ToothGrowth, facets = .~supp)
```
Conclusion: it looks like there are increases in tooth length by doses of OJ and VC. 

###Question 2: Summarizing the data by dose and supp.
```{r}
group_by(ToothGrowth, supp, dose) %>% summarize(mean = mean(len))
```
The above table summarizes the mean of tooth length for each dose/ supplement combination.

###Question 3: Compare dose and supplement delivery method using a t test
```{r}
OJdata_0.5 <-filter(ToothGrowth, supp=="OJ", dose == 0.5)
VCdata_0.5 <-filter(ToothGrowth, supp=="VC", dose == 0.5)
t.test(OJdata_0.5$len,VCdata_0.5$len)

OJdata_1.0 <-filter(ToothGrowth, supp=="OJ", dose == 1.0)
VCdata_1.0 <-filter(ToothGrowth, supp=="VC", dose == 1.0)
t.test(OJdata_1.0$len,VCdata_1.0$len)

OJdata_2.0 <-filter(ToothGrowth, supp=="OJ", dose == 2.0)
VCdata_2.0 <-filter(ToothGrowth, supp=="VC", dose == 2.0)
t.test(OJdata_2.0$len,VCdata_2.0$len)
```
It appears that there is a significant difference in tooth length when given low doses (0.5 and 1.0) of vitamin C through orange juice and through ascorbic acids, but this significance is no longer there at high doses (2.0)

### Question 4: Conclusion
Vitamin C probably promotes tooth growth in a dose-dependent manner between 0.5 and 2.0 milligram. At lower dosages, orange juice is a better delivery method than ascorbic acid. An assumption is that data from the two populations of guinea pig are independent and have identical distribution. 